---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

# Tree building using parsimony

The purpose of this short exercise is to show you how quick and simple it can be to generate a tree using R. 

We'll be using a morphological matrix with representatives of Ediacaran and Cambrian animals generated by the fab palaeobiologist [Francis Dunn](https://oumnh.ox.ac.uk/people/dr-frankie-dunn).
A link to the file we'll use for this exercise is below.
The original data set was downloaded from [here](https://research-information.bris.ac.uk/en/studentTheses/growth-and-development-in-the-ediacaran-macrobiota) and you can read the published results in [Science Advances](https://www.science.org/doi/abs/10.1126/sciadv.abe0291).
The purpose of building this tree was to try and establish the phylogenetic affinity of the enigmatic [*Charnia masoni*](https://en.wikipedia.org/wiki/Charnia), which is only known from the Ediacaran.

First install the necessary R packages. 


```{r packages, eval=FALSE}
install.packages("ape")
install.packages("phangorn")
```

`ape` is a very widely used phyologenetics package that contains (among many others) functions for reading, writing, and plotting phylogenetic trees and data.
`phangorn` is a package that contains functions for straightforward tree building. 

Download the character data from [here](data/Frankie-data.nex). Nexus is a common format used to store phylogenetic information. See if you can interpret the contents.
Next we'll read in the data and convert it into the format used by `phangorn`.

```{r data, eval=TRUE}
# read in the character matrix in ape 
dat <- ape::read.nexus.data(file = "data/Frankie-data.nex")

# identify character states
chars <- unique(unlist(unique(dat)))

# & convert to a phanghorn object
dat <- phangorn::phyDat(dat, type = "USER", levels = chars)
```

Next we need to provide `phangorn` with a starting tree.
We can do this using an approach called **neighbour-joining**. This simply calculates approximate evolutionary distance between the taxa in your matrix (using the `dist.hamming` function) and then groups taxa together based on their relative distances (using the `NJ` function).

```{r starting, eval=TRUE}
# create a neighbour joining starting tree
dm <- phangorn::dist.hamming(dat)
tree <- phangorn::NJ(dm)
```

Then we'll use the `phangorn` function `optim.parsimony` to find the most parsimonious tree.

```{r parsimony, eval=TRUE}
# estimate the parsimony tree
mp <- phangorn::optim.parsimony(tree, dat)
```

> How would you interpret the number output by this function?

Finally, we can specify an outgroup to root the tree. 
We'll use the fungus [*Laccaria*](https://en.wikipedia.org/wiki/Laccaria) because we know this taxa is more distantly related to all others in our tree than they are to each other.

```{r plot, eval=TRUE}
# identify the outgroup
outgroup <- "Laccaria"

# root the tree
mp <- ape::root(mp, outgroup = outgroup, resolve.root = TRUE)

plot(mp)
```

> Based on this tree how would you interpret the phylogenetic position of *Charnia*? Could you say whether it was a [crown group](https://en.wikipedia.org/wiki/Crown_group) or [stem group](https://en.wikipedia.org/wiki/Crown_group#Stem_groups) animal? 

Learn more about the taxa in this data set [here](latin_names.html).

*Charnia* was also the subject of this absolute [Palaeovision banger](https://www.youtube.com/watch?v=yeOC5CfSUxE).

## Other resources for building trees using parsimony

Because there are a large number of possible trees even for a small number of taxa, finding the most parsimonious tree is actually algorithmically really challenging. 
Many people use the open source software [TNT](http://www.lillo.org.ar/phylogeny/tnt/), which can speed up the tree search. 
You can also tweak your parsimony analysis based on assumptions you want to make about your character data. 
For example, characters can be ordered (e.g., you can allow character state changes from 0 to 1 or from 1 to 2 but prevent changes from 0 to 2). You can also assign different weights to different characters. 

We won't explore parsimony any further in this course, but it remains widely used for the analysis of morphological character data and is sometimes faster than alternative tree-building approaches. 
It can be a useful way to have an initial glimpse at your data/tree before or while running more computationally intensive analyses.

A complete script for this exercise can be downloaded [here](scripts/parsimony_example.R).

