---
title: ""
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

# Part 1

The first part of this course is a 3 day block course.
The aim is to introduce you to the major approaches used to reconstruct trees in paleobiology. By the end of the block course you should be able to reconstruct undated and dated trees using a Bayesian phylogenetic approach.

# Rough Schedule

* Day 1 - 10:00-16:00 (parsimony, likelihood, and substitution models)
* Day 2 - 10:00-16:00 (Bayesian inference, MCMC, clock models, node dating)
* Day 3 - 10:00-16:00 (the fossilised birth-death process)

The course will take place in the Henke Str. Pal seminar room. **You will need to bring your own laptop.**
Before course begins, make sure you have have [R](https://www.r-project.org) and [RStudio](https://posit.co/download/rstudio-desktop/) installed.

\

# Materials

Course material will be updated here throughout the week!

\

## Day 1

```{r, echo = FALSE}
#1. Course introduction and parsimony 14.04
#2. Likelihood and substitution models 14.04
#3. Bayesian inference and MCMC 15.04
#4. Clock models and node dating 15.04
#6. Dating using the fossilised birth-death process 16.04

library(kableExtra)

urls <- c("slides/01-intro.pdf","exercise-01.html",
          "slides/02-likelihood.pdf","exercise-02.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1", "Slides", "Exercise 2"),
                 topics = c("tree thinking, parsimony", "tree building using parsimony",
                            "maximum likelihood, substitution models", 
                            "tree building using maximum likelihood"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")
dt$schedule[4] <- paste0("[", dt$schedule[4], "](", urls[4], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* Re-familiarise yourself with [how to read a phylogenetic tree](https://artic.network/how-to-read-a-tree.html) and [rooting](https://slides.filogeneti.ca/include/rooting.html)
* Watch [A Brief History of Computational Phylogenetics](https://www.youtube.com/watch?v=PcD15i7yzJ8)

**Homework**

* Install the software [RevBayes](https://revbayes.github.io/download)
* Install the software [FigTree](http://tree.bio.ed.ac.uk/software/figtree/) and [Tracer](https://github.com/beast-dev/tracer/releases/tag/v1.7.2)
* Windows users should also install [Git Bash](https://gitforwindows.org)

\

## Day 2

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/03-Bayesian.pdf","exercise-03.html", "exercise-04.html")

dt <- data.frame(schedule = c("Slides", "Exercise 3", "Exercise 4"),
                 topics = c("Intro to RevBayes", 
                            "Intro to RevBayes and the Rev language", 
                            "Bayesian tree inference"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* [Bayesian Inference of Phylogeny and Its Impact on Evolutionary Biology](https://www.science.org/doi/epdf/10.1126/science.1065889)

\ 

## Day 3

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/04-clocks.pdf","exercise-05.html",
          "slides/05-FBD.pdf","exercise-06.html","exercise-06.html")

dt <- data.frame(schedule = c("Slides", "Exercise 5", "Slides", "Exercise 6", "Exercise 7"),
                 topics = c("Bayesian timetree inference", 
                            "Time tree inference in RevBayes (demo only)",
                            "The fossilized birth-death process",
                            "Simulations",
                           "The FBD process using RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")
dt$schedule[4] <- paste0("[", dt$schedule[4], "](", urls[4], ")")
dt$schedule[5] <- paste0("[", dt$schedule[5], "](", urls[5], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* [Bayesian molecular dating: opening up the black box](https://onlinelibrary.wiley.com/doi/abs/10.1111/brv.12390)
* [Understanding the tripartite approach to Bayesian divergence time estimation](https://ecoevorxiv.org/repository/view/4372/)

\
\

