---
title: ""
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

# Part 1

The first part of this course is a 3 day block course.
The aim is to introduce you to the major approaches used to reconstruct trees in paleobiology. By the end of the block course you should be able to reconstruct undated and dated trees using a Bayesian phylogenetic approach.

# Rough Schedule

* Day 1 - 12:00-16:00 (parsimony, likelihood, substitution models)
* Day 2 - 10:00-16:00 (Bayesian inference, MCMC, clock models, node dating)
* Day 3 - 10:00-16:00 (the fossilised birth-death process, total-evidence dating)

The course will take place in the Henke Str. Pal seminar room. **You will need to bring your own laptop.**
Before course begins, make sure you have have [R](https://www.r-project.org) and [RStudio](https://posit.co/download/rstudio-desktop/) installed.

\

# Materials

Course material will be updated here throughout the week!

\

## Day 1

```{r, echo = FALSE}
#1. Course introduction and parsimony 14.04
#2. Likelihood and substitution models 14.04
#3. Bayesian inference and MCMC 15.04
#4. Clock models and node dating 15.04
#6. Dating using the fossilised birth-death process 16.04

library(kableExtra)

urls <- c("slides/01-intro.pdf","exercise-01.html",
          "slides/02-likelihood.pdf","exercise-02.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1", "Slides", "Exercise 2"),
                 topics = c("tree thinking, parsimony", "tree building using parsimony",
                            "maximum likelihood, substitution models", 
                            "tree building using maximum likelihood"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")
dt$schedule[4] <- paste0("[", dt$schedule[4], "](", urls[4], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* Re-familiarise yourself with how to read a phylogenetic tree ([Part 1](https://artic.network/guides/how-to-read-a-tree), [Part 2](https://artic.network/guides/estimating-time-scales.html)) and [rooting](https://slides.filogeneti.ca/include/rooting.html)
* Watch [A Brief History of Computational Phylogenetics](https://www.youtube.com/watch?v=PcD15i7yzJ8)

**Homework**

* Install the software [MrBayes](https://nbisweden.github.io/MrBayes/)
* Install the software [BEAST2](http://beast2.org/)
* Install the software [FigTree](http://tree.bio.ed.ac.uk/software/figtree/) and [Tracer](https://github.com/beast-dev/tracer/releases/tag/v1.7.2)

\

## Day 2

```{r, echo = FALSE}

library(kableExtra)

#urls <- c("slides/03-Bayesian.pdf","exercise-03.html", "exercise-04.html")

dt <- data.frame(schedule = c("Slides", "Exercise 3", "Slides", "Exercise 4"),
                 topics = c("Bayesian tree inference", 
                            "tree building using MrBayes", 
                            "Intro to BEAST, molecular dating", "tree building using BEAST"))
#dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
#dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
#dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* [Bayesian Inference of Phylogeny and Its Impact on Evolutionary Biology](https://www.science.org/doi/epdf/10.1126/science.1065889)

\ 

## Day 3

```{r, echo = FALSE}

library(kableExtra)

#urls <- c("slides/04-clocks.pdf","exercise-05.html","slides/05-FBD.pdf","exercise-06.html","exercise-06.html")

dt <- data.frame(schedule = c("Slides", "Exercise 5", "Exercise 6", "Slides", "Exercise 7"),
                 topics = c("The fossilized birth-death process",
                            "Simulations", "The FBD process using BEAST2",
                            "Intro to morphological models",
                            "Total-evidence dating using BEAST2"))
#dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
#dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
#dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")
#dt$schedule[4] <- paste0("[", dt$schedule[4], "](", urls[4], ")")
#dt$schedule[5] <- paste0("[", dt$schedule[5], "](", urls[5], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Reading**

* [Bayesian molecular dating: opening up the black box](https://onlinelibrary.wiley.com/doi/abs/10.1111/brv.12390)
* [Understanding the tripartite approach to Bayesian divergence time estimation](https://ecoevorxiv.org/repository/view/4372/)

**Homework**

* Install the software [RevBayes](https://revbayes.github.io/download)
* Windows users should also install [Git Bash](https://gitforwindows.org)

\
\
