---
title: "Schedule"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The second part of the course takes place in the summer semester **Tuesdays 14:00-16:00 CET**.

\

# Provisional Schedule

```{r, echo = FALSE}

library(kableExtra)

dt <- data.frame(schedule = c("1", "2", 
                              "3", "4", 
                              "5", "6",
                              "7", "8",
                              "9", "10", "11", "12", "13"),
                 date = c( "14.04", 
                           "21.04", 
                           "28.04", 
                           "05.05", 
                           "12.05", 
                           "19.05", 
                           "26.05", 
                           "02.06", 
                           "09.06",
                           "16.06",
                           "23.06",
                           "30.06",
                           "07.06"),
                 topics = c("Total-evidence dating", 
                            "Diversification  rate estimation I",
                            "Diversification  rate estimation II",
                            "Introduction to PCMs", 
                            "**No class** (Paul Lewis lectures)", 
                            "**No class** (Paul Lewis lectures)", 
                            "Ancestral state reconstruction (homework)",
                            "Intro to RevBayes",
                            "**No class** (Project work)", 
                            "**No class** (Project work)",
                            "Model selection and model adequacy",
                            "Biogeography", 
                            "Project work"
                            ))

names(dt)[1] <- ""

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

\

# Materials

Course material will be updated here throughout the semester!

## Week 1

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/06-FBD-recap.pdf", "https://taming-the-beast.org/tutorials/FBD-tutorial/", "https://taming-the-beast.org/tutorials/Total-Evidence-Tutorial/")

dt <- data.frame(schedule = c("Slides", "Exercise 6", "Exercise 7"),
                 topics = c("Recap", "The FBD process (BEAST2)", "Total-evidence dating (BEAST2)"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

\

## Week 2

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/07-phylodynamics.pdf",  "https://taming-the-beast.org/tutorials/Total-Evidence-Tutorial/",
          "https://taming-the-beast.org/tutorials/MSBD-tutorial/")

dt <- data.frame(schedule = c("Slides", "Exercise 7", "Exercise 8"),
                 topics = c("Intro to phylodynamics", "Total-evidence dating (BEAST2)", "The MSBD package (BEAST2)"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

## Week 3

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/08-BDMMP-macro.pdf",  "https://taming-the-beast.org/tutorials/BDMM-Prime-for-macroevolution/")

dt <- data.frame(schedule = c("Slides", "Exercise 9"),
                 topics = c("Phylodynamics continued", "The BDMMPrime package (BEAST2)"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

## Week 4

```{r, echo = FALSE}

library(kableExtra)

urls <- c("slides/09-Intro-to-PCMs.pdf",  "https://nhcooper123.github.io/pcm-primer-online/cont.html")

dt <- data.frame(schedule = c("Slides", "Exercise 10"),
                 topics = c("Intro to PCMs", "Continuous trait models in R"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

## Weeks 5 and 6

During this scheduled break we'd like you to watch the following 4 video lectures by [**Paul Lewis**](https://plewis.github.io/) and answer a set of questions we have prepared. Paul Lewis has made numerous important contributions to the field of statistical phylogenetics, as well as producing excellent tools and materials for learning about phylogenetic theory.
The videos we'll use were developed as part of the [phyloseminar](http://phyloseminar.org) series. 

The videos will take about 3.5 hours to watch. Read the questions before you watch each video and write down your answers -- we'll go over these in week 6. **The goal is not to understand everything** but to strengthen your understanding of statistical phylogenetics.

Also make note of any questions you have after watching these and we'll try to answer them.

```{r, echo = FALSE}
library(kableExtra)

urls <- c("https://www.youtube.com/watch?v=1r4z0YJq580", "https://www.youtube.com/watch?v=UsLeY0wZr4Y", "https://www.youtube.com/watch?v=4PWlnNsfz90&t=45s", "https://www.youtube.com/watch?v=TLtOS--YwkU",
          "questions-01.html", "questions-02.html", "questions-03.html","questions-04.html", 
          "https://forms.gle/Y5wuBJwoiDQHp6jt5",
          "https://forms.gle/ixo73Gy4TTzcVoM96",
          "https://forms.gle/vrX6E65XhLukxo8W6",
          "https://forms.gle/SiA9JESnwcjw2eDm9")

dt <- data.frame(videos = c("Part 1", "Part 2", "Part 3a", "Part 3b"),
                 topics = c("trees and likelihood", "trees, likelihood and rate heterogeneity", "Intro to Bayesian phylogenetics", "Intro to Bayesian phylogenetics"),
                 questions = c("Questions 1", "Questions 2", "Questions 3", "Questions 4")) #forms = c("Form 1", "Form 2", "Form 3", "Form 4")

dt$videos[1] <- paste0("[", dt$videos[1], "](", urls[1], ")")
dt$questions[1] <- paste0("[", dt$questions[1], "](", urls[5], ")")
#dt$forms[1] <- paste0("[", dt$forms[1], "](", urls[9], ")")
dt$videos[2] <- paste0("[", dt$videos[2], "](", urls[2], ")")
dt$questions[2] <- paste0("[", dt$questions[2], "](", urls[6], ")")
#dt$forms[2] <- paste0("[", dt$forms[2], "](", urls[10], ")")
dt$videos[3] <- paste0("[", dt$videos[3], "](", urls[3], ")")
dt$questions[3] <- paste0("[", dt$questions[3], "](", urls[7], ")")
#dt$forms[3] <- paste0("[", dt$forms[3], "](", urls[11], ")")
dt$videos[4] <- paste0("[", dt$videos[4], "](", urls[4], ")")
dt$questions[4] <- paste0("[", dt$questions[4], "](", urls[8], ")")
#dt$forms[4] <- paste0("[", dt$forms[4], "](", urls[12], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

## Week 7

There is no class this week because we have the Berg! 🍺🥳 In this time you should complete the following tasks on ancestral state reconstruction using R. 

Start by reading [Section 3.2](https://nhcooper123.github.io/pcm-primer/foundations2.html#modelling-discrete-traits---simple-models) of *Phylogenetic Comparative Methods: A Pracitcal Introduction* by **[Natalie Cooper](https://www.nhm.ac.uk/our-science/people/natalie-cooper.html)**, including the sections on the Mk/ER, SYM, and ARD models. Next complete the [accompanying exercise](https://nhcooper123.github.io/pcm-primer-online/discrete.html#models-of-evolution-for-discrete-traits) in R.



```{r, echo = FALSE}

library(kableExtra)

urls <- c("https://nhcooper123.github.io/pcm-primer/foundations2.html#modelling-discrete-traits---simple-models",  "https://nhcooper123.github.io/pcm-primer-online/discrete.html#models-of-evolution-for-discrete-traits")

dt <- data.frame(schedule = c("Reading", "Exercise 11"),
                 topics = c("PCM Primer Section 3.2", "Ancestral state reconstruction in R"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


